Emotion Recognition in Written Conversations Using CASTLE

### 2021-01-14 msccls/MScTemplateCS.tex
Shit ton of changes

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{amsmath}

@@

%\geometry{showframe}
%better like this?
%\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
\students{Nick Persson}{dat14npe@student.lu.se}{Filip Karabeleski}{dat14fka.student.lu.se}

\thesisnumber{LU-CS-EX: 2020-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
% default is Master. Uncomment the following for "kandidatarbete"/Bachelor's thesis
%\thesistype{Bachelor}{Kandidatarbete}

%\title{Formatting a Master's Thesis}
\title{Keyword Categorization and Emotion Analysis in Voice Calls}

@@

\subtitle{Subtitle WIP}
\company{Telavox AB}
\supervisors
{Sara Lindgren, \href{mailto:Sara.Lindgren@telavox.com}{\texttt{Sara.Lindgren@telavox.com}}}
{Pierre Nugues, \href{mailto:Pierre.Nugues@cs.lth.se}{\texttt{Pierre.Nugues@cs.lth.se.com}}}
\examiner
{Flavius Gruian,\href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}

@@

\theabstract{
This document describes the Master's Thesis format for the theses carried out at 
the Department of Computer Science, Lund University.

Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words. It should be placed on a separate right-hand page, with an additional \textit{1cm} margin on both left and right. Avoid acronyms, footnotes, and references in the abstract if possible.

Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
}

\keywords{Machine Learning, NLP, Sentiment Analysis, Keyword Extraction, (style, structure)}

@@

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notations. Since natural languages continuously develop as time goes on and its adherence to structures diminish, it is difficult to define explicit rules to follow.

\subsection{Emotion Recognition}
Emotion has always been a key part of human language, and it therefore comes naturally that teaching a computer to understand emotion in human language is a big part of NLP. Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of piece of text as either positive or negative. Godtycklig bit om sentiment analysis. 
Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it. This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative. 
Emotion recognition has also been extended to work on different modalities, such as for example visual and auditory input.

Emotion recognition can be applied to data in various modalities, 
such as visual, in the form of facial expressions and body language, audible and text based in the form of dialogues. To o

The LSTM architecture \citep{ntlk2009} is something....

\begin{itemize}
    \item Förklara vad systemet är
    \item Berätta om företaget Telavox
\end{itemize}

\section{Purpose}
\begin{itemize}
    \item Varför behöver problemet lösas, nytta
\end{itemize}
\subsection{Telavox}
Telavox is a telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. 
\section{Research questions}
The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. \\
\textit{What level of precision can be reached with the categorization of keywords from voice calls?}\\
\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}\\
\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?}

\section{Terminology}
NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?)

Kan också kallas för Glossary and Abbreviations
\section{Delimitations}
Delimitations are choices made by the researcher which should be mentioned. They describe the boundaries that you have set for the study.
\section{Contributions}
Potentialen av applications 
\section{Related work}

@@

blablabla information om de olika fields of study
Deep learning is mostly successive neural networks but there are other types as well. Deep belief networks.

@@

\section{Machine Learning}
Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it. By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem. 
In \textbf{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number. The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.) 
The other main type of tasks in machine learning is \textbf{unsupervised learning}. The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector. The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.

@@

\begin{figure}[htp]
    \centering
    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
    \caption{An example of a single neuron}
    \label{fig:neuron}
\end{figure}

@@

Each of the layers in a deep learning network manipulate their input data in some way, and the result is stored in the layer's weights. The results are monitored with the help of the \textit{loss function}. The loss function returns a distance score that gives an indication on how far off the networks previous prediction was from the true values \citep{franoischollet2017learning}. The phenomena of reducing the distance score that the loss function produces is called training. 
Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.

\subsection{Gated Recurrent Unit - GRU}
A gated recurrent unit (GRU) is neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks. Basic GRU nodes are primarily constructed of an \textit{update gate} and a \textit{reset gate}. The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth. 
Although fairly new it is already a well-established variant of recurrent neural networks. In comparison to LSTM layers, GRU layers are cheaper to run since they use less parameters, but have less representational power. It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{msccls/explanatory_images/gru.png}
    \caption{A GRU node}
    \label{fig:gru_node}
\end{figure}

\subsection{Long Short-Term Memory - LSTM}
Another widely used neural network architecture is the Long Short-Term Memory (LSTM). It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem \citep{hochreiter1998}, that might occur in recurrent neural networks, and has barely been changed since. The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{msccls/explanatory_images/lstm.png}
    \caption{A LSTM node}
    \label{fig:lstm_node}
\end{figure}

\section{Natural Language Processing - NLP}
Känns vettigt att ha som section

\section{Word Vectors}
Word vectors are a type of word representation where texts are turned into numbers
\subsection{One-Hot Encoding}
The most basic way of creating a word vector is the one-hot encoding method. This creates a very big and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding. Each word is then represented by an array filled with zeroes except for one index set to one. For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the assignment of values to word:

@@

The other big issue with it is that no meaning can be extracted from the matrix. If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one hot vector, "cat" is as close to "dog" as it would be to "airplane".

\subsection{Word Embedding}
To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors. \citep{neuralnetworkmethods} 
Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word. According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use one-hot encoding. Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words. This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding. In addition to distance, direction could also be a useful tool to extract meaning from the embedding. %Temp for formatting, don't forget it's here

\begin{wrapfigure}{r}{0.25\textwidth}
\includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
\caption{Small example of a word embedding space}
\label{fig:direction}
\end{wrapfigure}

@@

\section{Evaluation}
There are many different metrics that can be used to evaluate a machine learning model, based on what kind of model and what its purpose is. For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}. In a confusion matrix, one of the axis represent the actual class of the instance the model is trying to predict and the other represents what instance the model predicted. The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}. In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}
    \caption{Binary Confusion Matrix}
    \label{fig:confusion}
\end{figure}

Whenever our model makes a prediction on an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed. 
For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative. The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class. For instance, if determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{n}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.

$$\frac{TP + TN}{TP + TN + FP + FN}$$

If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model. If, however, our set is imbalanced and contains e.g. 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless.

\subsection{Precision}
Precision is a metric that quantifies how many of the instances predicted positive that actually were true. It's calculated by taking all of the instances where the model predicted positive and was correct, and dividing it by all the instances where it predicted positive.

$$\frac{TP}{TP + FP}$$

>>>>

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{subfig}
\usepackage{textcomp}
\usepackage{calligra}

\newcommand{\icon}[1]{\includegraphics[height=30pt]{msccls/explanatory_images/CASTLE_logo.png}}

@@

%\geometry{showframe}
%better like this?
%\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
\students{Nick Persson}{nick@persson.dev}{Filip Karabeleski}{fkarabeleski@gmail.com}

\thesisnumber{LU-CS-EX: 2021-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
% default is Master. Uncomment the following for "kandidatarbete"/Bachelor's thesis
%\thesistype{Bachelor}{Kandidatarbete}

%\title{Formatting a Master's Thesis}
\title{Emotion Recognition in Written Conversations Using CASTLE}

@@

\subtitle{\protect\icon{CASTLE}}
\company{Telavox AB}
\supervisors
{Sara Lindgren, \href{mailto:Sara.Lindgren@telavox.com}{\texttt{Sara.Lindgren@telavox.com}}}
{Pierre Nugues, \href{mailto:Pierre.Nugues@cs.lth.se}{\texttt{Pierre.Nugues@cs.lth.se.com}}}
\examiner
{Flavius Gruian, \href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}

@@

\theabstract{
%This document describes the Master's Thesis format for the theses carried out at 
%the Department of Computer Science, Lund University.

%Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words. It should be placed on a separate right-hand page, with an additional \textit{1cm} margin on both left and right. Avoid acronyms, footnotes, and references in the abstract if possible.

%Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
}

\keywords{Machine Learning, NLP, Emotion Recognition}

@@

%ML är automagiskt

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation. Since natural languages naturally evolve as time goes on and their adherence to structures diminish, it is difficult to define explicit rules to follow.

%prata om emotion recognition

%The LSTM architecture \citep{ntlk2009} is something....

%\begin{itemize}
%    \item Förklara vad systemet är
%    \item Berätta om företaget Telavox
%\end{itemize}

\section{Purpose}
Customer service is centered around the relation between company and customer with the goal of pleasing the customer. This goal can be very difficult to achieve and there is no straightforward way to do so. By analyzing prior conversations it is possible for companies to gain valuable insights in the most effective way to please the customer. Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually. 
The company sponsoring this project approached us with an idea to try and solve this problem using machine learning. A usual technique for these types of problems is \textit{sentiment analysis} where utterances are labeled either as positive or negative. We instead decided to take another route, namely \textit{emotion recognition}. Emotion recognition focuses on labelling utterances from a non-binary set of emotions instead of sentiment analysis' binary method. This allows an analysis with greater specificity.   
In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.

\section{Telavox}
Telavox AB is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. As all of their customers deal in customer service they would all benefit from the solution previously discussed. Telavox therefore has a great interest in being able to offer this as a service. As Telavox has over 250 businesses as customers, the solution would facilitate the quality of customer service world wide.

\section{Research questions}
The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results. After this, a secondary aim is to investigate whether or not it is possible to create a dynamic model for instant response regarding customer service related conversations. The following questions are used as guidelines on how to realize these aims.

\begin{itemize}
    \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
    \item \textit{Is it possible for an LSTM architecture to produce better results than a GRU one for emotion recognition?}
    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
\end{itemize}

%NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?), utterance, speaker, dialogue,

%Hyperparameters en punkt i listan, items i hyperparameters en sublist

%Kan också kallas för Glossary and Abbreviations

\section{Contributions}
%Potentialen av applications 
\section{Related work}

@@

The celebrity of computer science at this time is undoubtedly artificial intelligence (AI). A much more complex subject than most people know. Most of the complex problems in AI are solved using technologies within its sub-fields, which are illustrated in Figure \ref{fig:map}, rather than traditional artificial intelligence such as robotics \citep{bajwa_2020}. 
In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model. Firstly, the model should be created with deep learning algorithms. This enables it to find patterns that it otherwise could not be able to. Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis.

@@

\section{Machine Learning}
Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it. By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem. 
In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number. The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.) 
The other main type of task in machine learning is \textit{unsupervised learning}. The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector. The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.

@@

%Weights control the signal (or the strength of the connection) between two neurons. In other words, a weight decides how much influence the input will have on the output.

\begin{figure}[htp]
    \centering
    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
    \caption{An example of a single neuron.}
    \label{fig:neuron}
\end{figure}

@@

Each of the layers in a deep learning network manipulates their input data in some way, and the result is stored in the layer's weights. The results are monitored with the help of the \textit{loss function}. The loss function returns a distance score that indicates how far off the network's previous prediction was from the true values \citep{franoischollet2017learning}. The process of reducing the distance score that the loss function produces is called training. 
Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.

\subsection{Recurrent Neural Network -- RNN}

One of the simpler neural network classes used in deep learning is recurrent neural networks (RNN) and they process either sequential data or time-series data \citep{lipton2015critical}. Since the inputs are sequential, all the inputs are dependant on each other in one way or another. RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.

One of RNN's weaknesses is that they suffer the \textit{vanishing gradient problem} \citep{hochreiter1998}. Theoretically, RNNs can process arbitrarily long sequences, in practice, they are limited to the length of the sequence. If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away. In a lot of cases, this will result in leaving out important information from the prediction. There are instead other alternatives that consider this weakness.

%Time series. Meant to process sequences of data like text, audio, video. Neural networks that feed their output back into their inputs recursively. One of the more basic building blocks in machine learning/deep learning. A gate looks like this:

%suffer from short-term memory. If a sequence is long enough, they’ll have a hard time carrying information from earlier time steps to later ones. So if you are trying to process a paragraph of text to do predictions, RNN’s may leave out important information from the beginning.

 \begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.3]{msccls/explanatory_images/rnn.png}
    \caption{A simple RNN node}
    \label{fig:rnn_node}
\end{figure}

\subsection{Gated Recurrent Unit -- GRU}
%Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
A gated recurrent unit (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks. Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}. The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth. 
Although fairly new it is already a well-established variant of recurrent neural networks. In comparison to LSTM layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power. It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.5]{msccls/explanatory_images/gru.png}
    \caption{A GRU node}
    \label{fig:gru_node}
\end{figure}

\subsection{Long Short-Term Memory -- LSTM}
Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM). It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since. The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[scale=0.45]{msccls/explanatory_images/lstm.png}
    \caption{A LSTM node}
    \label{fig:lstm_node}
\end{figure}

\section{Training Process}
In order to train a model there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input. When using a labeled data set the labels are regarded as the solution to the problem. When training a model the goal is to find values for its weights such that the model correctly labels the given input. The training of a model is done step by step in the shape of \textit{epochs}. Epochs represent the number of iterations that the model has passed through a data set intended for training the model. Following each epoch, a \textit{loss function} measures how bad the model's previous prediction was and produces a value for the \textit{loss}. The model will have a perfect prediction when the loss reaches zero, therefore minimizing the loss is desirable.

In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the

We want to travel through the function and find the place where the derivative is zero. This is done for each weight in the network. It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero. Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.

direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.

%In gradient descent one is trying to reach the minimum of the loss function with respect to the parameters using the derivatives calculated in the back-propagation.

%gradient descent iteratively tries to find a minimum. tries to regulate the weights in order to minimize loss.

%Datasets ger inputs till model, första lagret ..... Modellen har svaren till hands och använder dessa för att styra skeppet i rätt riktning. 
%There exist many components to training a model .....
%In \ref{chapter}, backpropagation and gradient descnet among other temrs were mentioned but not extensively explored. The training process can be described as the road from a standalone model to having relevant results.

% In this context, learning means finding a set of values for the weights of all layers in a network, such that the network will correctly map example inputs to their associated targets

%Backpropagation
%Gradient Descent
%Loss (function)
%Från början till slut, från en modell med lager till en evaluation, sen fortsätter vi snacka om det under evaluation
%Overfitting
%Epochs

\section{Evaluation}
Many different metrics can be used to evaluate a machine learning model, based on what kind of model and what its purpose is. For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}. In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted. The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}. In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}

\begin{figure}[h!]
    \centering
    \hbox{\hspace{6em}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}}
    \caption{Binary Confusion Matrix}
    \label{fig:confusion}
\end{figure}

Whenever our model predicts an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed.

For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative. The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class. For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.

$$ Accuracy = \frac{TP + TN}{TP + TN + FP + FN}$$

If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model. If, however, our set is imbalanced and contains e.g. 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless \citep{imbalancedlearning}.

\subsection{Precision}
Precision is a metric that quantifies how many of the instances predicted positive that were true. It's calculated by taking all of the instances where the model predicted positive and was correct and dividing it by all the instances where it predicted positive.

$$ Precision = \frac{TP}{TP + FP}$$

Precision gives us information on how accurate the model is when predicting an instance to be true. This is useful when dealing with instances where it's important to not give a false positive. If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal. 
However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.

\subsection{Recall}
Recall is a metric that quantifies how many percent of positive instances that our model managed to predict which we calculate by dividing all instances the model predicted positive and was right by all positive instances.

$$ Recall =\frac{TP}{TP + FN}$$

As a sort of counterpoint to precision which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed. A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating. Recall doesn't take into account how many positive predictions turned out to be false so in cases where that doesn't matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.

\subsection{F1-score}
A combination of the two previous metrics, F1-score is calculated by creating the harmonic mean of \textit{precision} and \textit{recall}.

$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall} $$

In cases where there is no extra importance in finding either false negatives or false positives, you just want to minimize the risk of errors whatsoever. The harmonic mean is a kind of average that is appropriate when working with ratios, which is what both precision and recall are. This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision. In multi-class scenarios with imbalanced data sets there are multiple ways to calculate the F1-score, namely \textit{macro}, \textit{weighted} and \textit{micro}. Which one to use depends on your data set and what you prioritize in your model.

The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to disproportionate weight on the classes with fewer instances. If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50. The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones. When calculating the mean F1-score, the low fear score would muddy the entire score and make it seem bad, when in fact the score for the neutral case is excellent. If we want to optimize for the minority classes, however, this lets us do just that.

To counteract this behavior of favoring the minority class, one could instead calculate the weighted F1-score. After calculating the F1-score for each class, they are weighted by the number of true instances for each class, thus making them proportionate before calculating the final score. This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which might not be the goal of the model.

When calculating micro F1-score, instead of calculating the f1-scores separately, the f1-score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class. If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class. From the perspective of the joy class, however, it would instead be an FP. Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases. Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct. This means that just like Accuracy this is a good metric on balanced data sets but can be deceiving in an imbalanced one.

\section{Loss}
%might be removed

\section{Overfitting}
Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set. Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set. This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.

\section{Natural Language Processing -- NLP}
Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy like e.g English or Mandarin. Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}. Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game} or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human. One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm. Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.

%is used for classification, translation, emotion/sentiment analysis,

Emotion is a key part of human language and it, therefore, comes naturally that teaching a computer to understand emotion in human language is a big part of NLP. Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of a piece of text as either positive or negative. 
%Godtycklig bit om sentiment analysis. 
Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it. This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative. 
Emotion recognition has also been extended to work on different modalities, such as visual and auditory input.

%Emotion recognition can be applied to data in various modalities, 
%such as visual, in the form of facial expressions and body language, audible and text-based in the form of dialogues.

\subsection{Word Vectors}
%Om vi behöver en tie in: Instead of analysing the words themselves, they are converted into numerical representations that are easier for the computer to handle.

Word vectors are a type of word representation where texts are turned into a series of numbers, each word or sometimes phrase instead represented by a number.

\subsection{One-Hot Encoding}
The most basic way of creating a word vector is the one-hot encoding method. This creates a very big and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding. Each word is then represented by an array filled with zeroes except for one index set to one. For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the following assignment of values to words:

@@

The other big issue with it is that no meaning can be extracted from the matrix. If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one-hot vector, "cat" is as close to "dog" as it would be to "airplane".

\subsection{Word Embedding}
To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors. \citep{neuralnetworkmethods} 
Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word. According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional, or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use a one-hot encoding. Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words. This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding. In addition to distance, the direction could also be a useful tool to extract meaning from the embedding. %Temp for formatting, don't forget it's here

\begin{wrapfigure}{r}{0.25\textwidth}
\includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
\caption{Small example of a word embedding space. From \citet{franoischollet2017learning}.}
\label{fig:direction}
\end{wrapfigure}

@@

%Pytorch
%GRU
%Context-based 
%Commonsense Features
%Förklara bilden

%\section{Theory}
%NLP technique, categorization, ML, supervised/unsupervised,  
%Mechanical turks to 
%Finns både text och audio på MELD, \citep{zhang2019modeling} visar att text för sig är bättre än ljud men att text+ljud > text
%train, dev(validation), test, man testar på dev under development, först på test efter man är klar

\chapter{Implementation}
%Summary of what the chapter is about
The following chapter is mainly focused on giving a clear description of what we have done for the duration of the project. Firstly we will describe our model \textit{CASTLE}, how it has been implemented and what changes were made from the baseline \textit{COSMIC}. Our initial goal was to have our model produce state-of-the-art results in emotion recognition which later on changed to a more primary focus on optimizing the run time instead. 
We based our research strategy on the work by \citet{bouthillier:hal-02447823} that summarizes and reviews the experimental procedures that are most commonly used in the field of machine learning.

\section{Baseline -- COSMIC}
When trying to improve existing results in many machine learning fields, instead of starting from scratch, a baseline is used. A baseline is an already functioning model that provides basic results to the problem at hand. This provides the opportunity to allocate resources on improving the model instead of building it from the ground up and allows the machine learning community to expand on and draw benefits from each other's work.

This project has been developed using the COSMIC model \citep{ghosal2020cosmic} as a baseline. COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four data sets commonly used in emotion recognition. The idea behind the model is when classifying an utterance, the emotions of the utterances both preceding and succeeding it are also taken into account. This way the context of the conversation can be used to more accurately predict the emotions of individual utterances.

%COMSICs arkitektur bygger på att de skapar egna GRU-celler och GRU-lager som bygger på PyTorchs respektive. 
%Filip

Figure \ref{fig:COSMIC_arch} provides a simplified description of how COSMIC operates in a conversation between person A and person B. For each time step $t$ there is an utterance said by either A or B with the other person as a listener. At the next time step $t+1$ the roles are alternated. For each utterance, there exist multiple GRU-cells that predict both the intent of the speaker and the response of the listener at a certain time step. These cells are used not only to predict the emotion of the current utterance but also used for the prediction of emotions in other utterances related to said utterance.
